### 2022-02-09 chapters/intro/main.tex
modifs avant le sik

The latest major outbreak affecting coral reefs of the Caribbean is the SCTLD \citep{noaa2018}. First observed off the coasts of Miami in 2014 during the PMDDP \citep{precht2016unprecedented}, the disease has since spread throughout the entire Florida Reef Tract and numerous territories of the Caribbean \citep{alvarez2019rapid, kramer2019map, estrada2021effects}. The outbreak affects at least 24 species of scleractinian coral and often results in whole colony mortality \citep{precht2016unprecedented, walton2018impacts}. Briefly, the gross morphology of SCTLD can be focal or multifocal, with locally extensive to diffuse areas of acute to subacute tissue loss distributed basally, peripherally, or both. In some cases, tissues bordering areas of chronic tissue loss show indistinct bands (1–5 cm) of pallor, progressing to normal pigmentation away from the denuded skeleton. The continued persistence of the outbreak, the high number of species affected, and its large geographical distribution of reports consistent with the case definition suggests that SCTLD is the largest coral disease outbreak on record in Florida. Although the causative agent of the disease remains unknown, hydrodynamics are likely to play an important role in its propagation as both modeling studies and ex situ experiments show evidence of waterborne disease transmission \citep{aeby2019pathogenesis,dobbelaere2020coupled,eaton2021measuring,meiling2021variable}. Furthermore, recent studies showed evidence that sediments can act as a vector for the SCTLD \citep{rosales2020rhodobacterales, studivan2022reef}.

@@

Additionally, Florida is a prime landfall target for tropical cyclones from June through November (Fig. \ref{inro:landfall}). Hurricanes that forms in June and July are usually weak while hurricanes occurring in August and September tend to become severe storms \citep{banks2008reef}. This is illustrated by Hurricane Irma, one of the strongest and costliest hurricanes on record in the Atlantic, that made landfall in Florida in September 2017 as a category 4 hurricane \citep{irmaNOAA, xian2018brief}. Hurricanes are a major agent of coral mortality in the Caribbean that contributed to teh decline of acroporids \citep{gardner2003long,aronson2001white}. They cause a wholesale destruction of the reef biota on their passage and can induced coral burial through sediment resuspension \citep{banks2008reef, miller2008effects}. Furthermore, they significantly impact the content of surface waters through upwelling and mixing, which can disturb the benthic communities \citep{wachnicka2019hurricane,varlas2020investigating}. Furthermore, these

@@

\section{Modeling the hydrodynamics of the FRT}

>>>>

The latest major outbreak affecting coral reefs of the Caribbean is the SCTLD \citep{noaa2018}. First observed off the coasts of Miami in 2014 during the PMDDP \citep{precht2016unprecedented}, the disease has since spread throughout the entire Florida Reef Tract and numerous territories of the Caribbean \citep{alvarez2019rapid, kramer2019map, estrada2021effects}. The outbreak affects at least 24 species of scleractinian coral and often results in whole colony mortality \citep{precht2016unprecedented, walton2018impacts}. Briefly, the gross morphology of SCTLD can be focal or multifocal, with locally extensive to diffuse areas of acute to subacute tissue loss distributed basally, peripherally, or both. In some cases, tissues bordering areas of chronic tissue loss show indistinct bands (1–5 cm) of pallor, progressing to normal pigmentation away from the denuded skeleton. The continued persistence of the outbreak, the high number of species affected, and its large geographical distribution of reports consistent with the case definition suggests that SCTLD is the largest coral disease outbreak on record in Florida. Although the causative agent of the disease remains unknown, there is evidence of waterborne disease transmission \citep{aeby2019pathogenesis,eaton2021measuring,meiling2021variable} and recent studies showed evidence that sediments can act as a vector for the SCTLD \citep{rosales2020rhodobacterales, studivan2022reef}.

@@

Additionally, Florida is a prime landfall target for tropical cyclones from June through November (Fig. \ref{inro:landfall}). Between 1899 and 1998, Southeastern Florida has been directly hit by 27 hurricanes, 5 of which were of category 4 or more (winds > 209 km/h) \citep{neumann1999tropical}. Hurricanes that forms in June and July are usually weak while hurricanes occurring in August and September tend to become severe storms \citep{banks2008reef}. This is illustrated by Hurricane Irma, one of the strongest and costliest hurricanes on record in the Atlantic, that made landfall in Florida in September 2017 as a category 4 hurricane \citep{irmaNOAA, xian2018brief}. Hurricanes are a major agent of coral mortality in the Caribbean that contributed to teh decline of acroporids \citep{gardner2003long,aronson2001white}. They cause a wholesale destruction of the reef biota on their passage and can induced coral burial through sediment resuspension \citep{banks2008reef, miller2008effects}. Furthermore, they significantly impact the content of surface waters through upwelling and mixing, which can disturb the benthic communities \citep{wachnicka2019hurricane,varlas2020investigating}. Furthermore, heavy wind conditions generate important waves that interact with currents and significantly alter transport processes in nearshore waters \citep{niu2017role,mao2020particle}. This might impact the dispersion of coral larvae, that are released during the hurricane season \citep{frys2020fine}. Hurricane might therefore interrupt larval exchanges between otherwise connected reef but also create new connectivity pathways. As major hurricanes are intensifying under the effect global warming \citep{bhatia2019recent,knutson2020tropical}, better understanding the impact of hurricane on wave-induced processes could inform the management of reefs under future climatic conditions. Furthermore, as incoming storm-generated waves dissipate over reefs, better understanding the protective role of coral reefs might promote their restoration as tool to reduce coastal flooding hazard \citep{roelvink2021coral}.

@@

This thesis focuses on two main threats influencing the future and survival of FCR: the current outbreak of SCTLD and the hurricanes. More specifically, our goal is to develop modeling tools to answer four main questions: (1) how does the SCTLD spread from reef to reef ? (2) how can we explain the timing of this spread ? (3)

Answering these questions require the development

\section*{Modeling the hydrodynamics of the FRT}
